---
title: "Col I/CNA35 staining_image analyis_initial steps"
author: "Nadine Tuechler"
date: "**Date**: `r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
  word_document: default
editor_options: 
  chunk_output_type: console
---

This script was generated with help from Denes Turei and Sarah Kaspar

# Data

- Count_NucleiAll: indicates the nuclei (cell) number per image  
- Col1_bg_corrected
    - We want to measure the Collagen 1 deposition that is by treating with TGF  
    - Background of the cover slip is already subtracted with: Col1_bg_corrected = Intensity_TotalIntensity_Col1 - (MEAN_background * TOTAL_AREA)  
    - MEAN_background is 200/2^16  
    - TOTAL_AREA = 4665600 (pixel number)  
    - No autofluorescence correction yet (this is what we use the condition “unstained” for)  

- plate: biological replicate, 24-well plate  
- well: wells within the plate  
- position: position within a well (36 positions per well)  
- time: 0h, 12h, 24h, 48h, 72h, 96h  
    - Time of treatment with TGF or the ctrl  
    - 0h ctrl is NOT like the other controls  
- condition: ctrl, TGF, unstained  
- replicates: technical replicates on a plate  


# 1. Load libraries and data
```{r load libraries}
#BiocManager::install("dplyr")

#load libraries
library(tidyverse)
library(fitdistrplus)
library(vcd)
library(readr)
library(dplyr)
library(magrittr)
library(stringr)
library(tidyr)
```

```{r parameters}
PARAM <- list()
PARAM$folder <- paste0(getwd(), "/")
PARAM$folder.R <- gsub("code/", "",PARAM$folder)
PARAM$folder.output <- paste0(PARAM$folder.R, "output/")
PARAM$folder.figures <- paste0(PARAM$folder.R, "figures/linear_model/")
PARAM$folder.data <- paste0(PARAM$folder.R, "data/")

```

```{r import data and get rid of unnecessary rows}
import_files <- function(filename){
  df <-   file.path(PARAM$folder.data, filename) %>% 
  {suppressWarnings(readLines(.))} %>%
  {c(.[1], stringr::str_subset(., '^\\d'))} %>%
  paste(collapse = '\n') %>%
  readr::read_csv(col_types = readr::cols())
}

df_A <- import_files(filename = 'NT22_004_A_20240127.csv')
df_B <- import_files(filename = 'NT22_004_B_20240127.csv')
df_C <- import_files(filename = "NT22_004_C_20240127.csv")
df_D <- import_files(filename = "NT22_004_D_20240127.csv")


plotdir <- PARAM$folder.figures <- paste0(PARAM$folder.R, "figures/linear_model/")
```


# 2. Optimize dataframe
extract information like well name and position
```{r df extract well name and position}
# rep A
df_A %<>%
  tidyr::extract(
    FileName_Collagen,
    c('well', 'position'),
    'Col1-A--(\\w\\d)[\\w\\d-]+(P\\d+)'
  )
df_A$plate <- "A"

# rep 
df_B %<>%
  tidyr::extract(
    FileName_Collagen,
    c('well', 'position'),
    'Col1-B--(\\w\\d)[\\w\\d-]+(P\\d+)'
  )
df_B$plate <- "B"

# rep C of NT23_012
df_C %<>%
  tidyr::extract(
    FileName_Collagen,
    c('well', 'position'),
    'Col1-C--(\\w\\d)[\\w\\d-]+(P\\d+)'
  )
df_C$plate <- "C"

# rep D
df_D %<>%
  tidyr::extract(
    FileName_Collagen,
    c('well', 'position'),
    'Col1-D--(\\w\\d)[\\w\\d-]+(P\\d+)'
  )
df_D$plate <- "D"
```


# 3. Add treatments --> new dataframe 
```{r df define treatments}
# treatment dfs
treatments_A <-
  file.path(PARAM$folder.data, 'NT22_004_treatments_A.csv') %>%
  readr::read_csv(col_types = readr::cols())

treatments <-
  file.path(PARAM$folder.data, 'NT22_004_treatments.csv') %>%
  readr::read_csv(col_types = readr::cols())
```

get separate column with treatments
```{r df include treatments}
# first set of siRNAs
df_A %<>%
  dplyr::inner_join(treatments_A, by = 'well') %>%
  tidyr::separate(., treatments, c('time', 'condition'), fill = 'left')

df_B %<>%
  dplyr::inner_join(treatments, by = 'well') %>%
  tidyr::separate(., treatments, c('time', 'condition'), fill = 'left')

df_C %<>%
  dplyr::inner_join(treatments, by = 'well') %>%
  tidyr::separate(., treatments, c('time', 'condition'), fill = 'left')

df_D %<>%
  dplyr::inner_join(treatments, by = 'well') %>%
  tidyr::separate(., treatments, c('time', 'condition'), fill = 'left')

```

# 4. multiply values wit 65535 
to have the same values that are shown in image J
```{r}
# multiply with 65535 
# write function to multiply values with 65535 (as shown in image J)

multiply65535 <- function(df){df %>% 
  dplyr::mutate(Intensity_TotalIntensity_Collagen_x65535 = Intensity_TotalIntensity_Collagen * 65535)}

# multiply 

df_65535_A <- multiply65535(df_A)
df_65535_B <- multiply65535(df_B)
df_65535_C <- multiply65535(df_C) 
df_65535_D <- multiply65535(df_D)

```


# 5. Background subtraction
## function bg_subtract
```{r}
# write function to subtract background and unspecific antibody staining
bg_subtract <- function(df, MEAN_BG_65535, MEAN_BG_ABunspecific){
  
  # total area of image 
  TOTAL_AREA <- 4665600 

  # cover slip background subtraction
  df %>%
    dplyr::mutate(
      Col1_bg_corrected_65535 = Intensity_TotalIntensity_Collagen_x65535 - (MEAN_BG_65535 * TOTAL_AREA)) %>%

  # unspecific antibody staining background subtraction 
  # (not for unstained control as there is no primary antibody)
   dplyr::mutate(
      Col1_bg_and_unspecificAB_corrected_65535 = ifelse(condition !="unstained",
        Col1_bg_corrected_65535 - ((MEAN_BG_ABunspecific * (TOTAL_AREA - (0 * Count_Nuclei)))), #if
        Col1_bg_corrected_65535) #else
      ) %>%
  
    # normalize to nuclei number 
    dplyr::mutate(intensity_diff_nNuclei_65535 = Col1_bg_and_unspecificAB_corrected_65535 / Count_Nuclei)

  #%>%
   #dplyr::mutate(cell_free_area = TOTAL_AREA - (0 * Count_Nuclei))
     #    Col1_bg_corrected_65535_2 = Col1_bg_corrected_65535 - (MEAN_BG_ABunspecific * TOTAL_AREA))

}

```


### subtract backgrounds
```{r}
df_65535_new_A <- bg_subtract(df_65535_A, MEAN_BG_65535 = 220, MEAN_BG_ABunspecific = 50)
df_65535_new_B <- bg_subtract(df_65535_B, MEAN_BG_65535 = 220, MEAN_BG_ABunspecific = 55)
df_65535_new_C <- bg_subtract(df_65535_C, MEAN_BG_65535 = 234, MEAN_BG_ABunspecific = 65)
df_65535_new_D <- bg_subtract(df_65535_D, MEAN_BG_65535 = 232, MEAN_BG_ABunspecific = 80)

```


example to check background subtraction of unstained control
```{r}
plot_Col1_corr_per_cell_unstained_A <-
  ggplot(data = df_65535_new_A %>% 
           filter(condition == "unstained"),
         aes(x = Count_Nuclei, y = intensity_diff_nNuclei_65535, colour = condition, shape = replicates)) + # y = Col1_per_cell
  geom_point() + 
#  facet_wrap( ~ time) +
  xlim(0,450)+
  ggtitle("Col1 (BG,AF) corrected per cell - plate A") #+
#  ylim(-5000000,15000000)
  
plot_Col1_corr_per_cell_unstained_A
```


## 6. Auto-fluorescence correction 
### Mean Auto-fluorescence (unstained) intensity per nucleus
function for auto-fluorescence correction
```{r}
AF_correction <- function(df){
  df %<>%
    # calculate fluorescence per cell
    filter(condition == "unstained") %>%
    subset(Count_Nuclei > 20) %>%

   dplyr::mutate(AFcell = mean(intensity_diff_nNuclei_65535))
}

```


function for auto-fluorescence subtraction 
```{r}
## Normalize to nuclei number and subtract autofluorescence per cell 
AF_subtraction <- function(df_first_AF, df_second){
    
    # auto-fluorescence signal
    AF_per_cell <- unique(df_first_AF$AFcell) 
  
    # norm to nuclei and subtract AF
    df_second %<>%
      filter(condition != "unstained") %>%
 
    # subtract auto-fluorescence signal per cell
    mutate(Col1_per_cell = intensity_diff_nNuclei_65535 - AF_per_cell)

}

```

```{r}
# calculate autofluorescence signal 
df_A_AF <- AF_correction(df_65535_new_A)
df_B_AF <- AF_correction(df_65535_new_B)
df_C_AF <- AF_correction(df_65535_new_C)
df_D_AF <- AF_correction(df_65535_new_D)

# subtract autofluorescence per cell
#df_A12_final <- AF_subtraction(df_first_AF = df_A12_AF, df_second = df_65535_A)
df_A_final <- AF_subtraction(df_A_AF, df_65535_new_A)
df_B_final <- AF_subtraction(df_B_AF, df_65535_new_B)
df_C_final <- AF_subtraction(df_C_AF, df_65535_new_C)
df_D_final <- AF_subtraction(df_D_AF, df_65535_new_D)

```


```{r}
plot_Col1_corr_per_cell_AA <-
  ggplot(data = df_A_final, 
         aes(x = Count_Nuclei, y = Col1_per_cell, colour = condition, shape = replicates)) + # y = Col1_per_cell
  geom_point() + 
  facet_wrap( ~ time) +
  #xlim(0,450)+
  ggtitle("Col1 (BG,AF) corrected per cell - plate A") # +
#  ylim(-5000000,150000000)
  
plot_Col1_corr_per_cell_AA
```

two outliers at 48h  --> remove values above 1.2e+07 (but should be gone with >20 nuclei filter 

```{r}
# show which ones are the outlier images 
outlier <- which(df_A_final$Count_Nuclei < 20) # & 
                    #df$Count_NucleiAll<350 )#returns number of lines to be selected
df_outliers <- df_A_final[outlier , ]


# df with data that do not include the outliers
outlier_ex <- which(df_A_final$Count_Nuclei > 20) # & 
                    #df$Count_NucleiAll<350 )#returns number of lines to be selected
df_outliersEX <- df_A_final[outlier_ex , ]
#data <- df_outliersEX
```

```{r}
plot_Col1_corr_per_cell_A <-
  ggplot(data = df_outliersEX, 
         aes(x = Count_Nuclei, y = Col1_per_cell, colour = condition, shape = replicates)) +
  scale_color_manual(values = c("#424242", "#757575")) +
  geom_point(alpha = 0.6) + 
  facet_wrap( ~ time) +
  xlim(0,450)+
  ggtitle("Col1 (BG,AF) corrected per cell - plate A") +
  ylim(-500000,12000000)
  
plot_Col1_corr_per_cell_A
```

does collagen deposition increase over time in ctrl?? -- might have to adjust background correction 



# 7. Create full df with all replicates
```{r create full df}
data <- rbind(df_A, df_B, df_C, df_D)
#data %<>% filter(siRNA != "NA")
dim(data)
```


```{r}
plot_Col1_corr_per_cell_all <-
  ggplot(data = data, # %>%
           #subset(Count_Nuclei >20),
         aes(x = Count_Nuclei, y = Col1_per_cell, colour = condition, shape = replicates)) + 
  scale_color_manual(values = c("#424242", "#757575")) +
  geom_point(alpha = 0.6) + 
  facet_grid(plate ~ time) +
  xlim(0,450)+
  ggtitle("Col1 (BG,AF) corrected per cell - all") #+
  ylim(-5000000,150000000)
  
plot_Col1_corr_per_cell_all
```


## exclude images with less than e.g. 20 nuclei 

```{r}
# show which ones are the outlier images 
outlier <- which(data$Count_Nuclei < 20, data$Col1_per_cell > 12000000) # & 
                    #df$Count_NucleiAll<350 )#returns number of lines to be selected
df_outliers <- data[outlier , ]


# df with data that do not include the outliers
outlier_ex <- which(data$Count_Nuclei > 20)
                    #df$Count_NucleiAll<350 )#returns number of lines to be selected
df_outliersEX <- data[outlier_ex , ]
#data <- df_outliersEX
```

```{r}
plot_Col1_corr_per_cell_all <-
  ggplot(data = df_outliersEX,
         aes(x = Count_Nuclei, y = Col1_per_cell, colour = condition)) + # y = Col1_per_cell
    scale_color_manual(values = c("ctrl" = "#4F87BF", 
                                "TGF" = "#80146E")) +
  geom_point(alpha = 0.6, size= 0.8) +
  facet_grid(plate ~ time) +
  xlim(0,450)+
  ggtitle("Col I (BG,AF) corrected per cell - Count_Nuclei > 20") +
  ylim(-500000,30000000)+
  theme_bw()
  
plot_Col1_corr_per_cell_all

ggplot2::ggsave(filename = paste0(PARAM$folder.figures,
                                  Sys.Date(), "_", "NT22_004_all.png"),
                plot = plot_Col1_corr_per_cell_all +
                  theme(axis.text = element_text(size = 9),
                        axis.text.x = element_text(size=9, angle = 45, vjust = 1, hjust = 1),
                        strip.text = element_text(size = 11),
                        axis.title = element_text(size = 11),
                        legend.text = element_text(size = 9),  # Adjust size of legend text
                        legend.title = element_text(size = 11)),
                width = 8, height = 5)
```

# save data
```{r}
dir.create('output', showWarnings = FALSE)


df_outliersEXnew <- subset(df_outliersEX, select=-c(Metadata_FileLocation))

file.path('output', 'NT22_004_Col1_20240127.csv') %>%
readr::write_csv(file = ., df_outliersEXnew)
```

to find images to show in the thesis 
```{r}
# means <- df_outliersEXnew %>%
#   group_by(plate, time, condition) %>%
#   summarise(mean = mean(Col1_per_cell))
```



# 8. Calculate ratios
ratio of TGF vs ctrl 

## ratio ctrl vs TGF

create df with intensity_norm_nuclei number values for each TGF treated sample
```{r}
ctrl_df <- df_outliersEXnew %>%
  filter(condition == "ctrl") %>%
  group_by(plate, time) %>%
  summarise(ctrl_mean_intensity = mean(Col1_per_cell))
```

### ratio TGF
```{r}
df_ratio <- df_outliersEXnew %>%
  #filter(condition == "TGF") %>%
  left_join(ctrl_df, by = c("plate", "time")) %>%
  mutate(FC_intensity = Col1_per_cell / ctrl_mean_intensity) %>%
  group_by(time, condition)%>%
  mutate(mean_FC = mean(FC_intensity))

```

#### plot
```{r}
# p_test <- df_ratio %>%
#  #filter(condition == "TGF") %>%
#   ggplot(aes(x = condition, y = FC_intensity)) +
#   stat_summary(geom = "crossbar", fun= median, color = "black")+
#   ggpubr::stat_compare_means(ref.group = "ctrl", label = "..p.signif..", method = 't.test',hide.ns = T, size = 5, label.y = 4.8, show.legend = F)+# wilcoxon
#   ggforce::geom_sina(aes(color = condition, shape = plate),size = 0.8, alpha = 0.6)+
#     scale_color_manual(values = c("ctrl" = "#4F87BF", 
#                                 "TGF" = "#80146E")) +  facet_wrap(~time)+
#   geom_hline(yintercept = 1, color = "gray26", linetype = "longdash") + #C0C0C0 708090 B2BEB5, B0C4DE
#   ggtitle("FC = Col1_per_cell_TGF / mean_intensity_ctrl") +
#   #ylim(0,5)+
#   theme_bw()
#   
# p_test
# 
# ggplot2::ggsave(filename = paste0(PARAM$folder.figures,
#                                   Sys.Date(), "_", "NT22_004_TGFvsctrl.pdf"),
#                 plot = p_test +
#                   theme(strip.text = element_text(size = 10)),#+
#                   #theme(axis.text = element_text(size = 11)),
#                 width = 7, height = 5)
```

## ratio ctrl vs TGF

```{r}
df_ratio_2 <- df_outliersEXnew %>%
  #filter(condition == "TGF") %>%
  left_join(ctrl_df, by = c("plate", "time")) %>%
  mutate(FC_intensity_2 = (Col1_per_cell - ctrl_mean_intensity) / ctrl_mean_intensity)

```

#### plot
```{r}
p_test_2 <- df_ratio_2 %>%
 #filter(condition == "TGF") %>%
  ggplot(aes(x = condition, y = FC_intensity_2)) +
  stat_summary(geom = "crossbar", fun= median, color = "black")+
  ggpubr::stat_compare_means(ref.group = "ctrl", label = "..p.signif..", method = 't.test')+# wilcoxon
  ggforce::geom_sina(aes(color = condition, shape = plate),size = 0.8, alpha = 0.6)+
  scale_color_manual(values = c("#424242","#9E9E9E")) +#, #008B8Bggforce
                                # "A" = "#80146E", #2F4F4F
                                # "B" = "#7958AC", #00CED1
                                # "C" = "#2BACCA"
 # geom_jitter(aes(color = plate),size = 1, width = 0.25, alpha = 0.6)+
  facet_wrap(~time)+
  geom_hline(yintercept = 0, color = "gray26", linetype = "longdash") + #C0C0C0 708090 B2BEB5, B0C4DE
  ggtitle("LogFC TGF vs ctrl per time") +
  #ylim(0,5)+
  theme_bw()
  
p_test_2

ggplot2::ggsave(filename = paste0(PARAM$folder.figures,
                                  Sys.Date(), "_", "NT22_004_TGFvsctrl_2.pdf"),
                plot = p_test_2 ,#+
                  #theme(axis.text = element_text(size = 11)),
                width = 7, height = 5)
```

# compare to 0h

create df with intensity_norm_nuclei number values for each TGF treated sample
```{r}
ctrl_df_0h <- df_outliersEXnew %>%
  filter(condition == "ctrl" & time == "0h") %>%
  group_by(plate) %>%
  summarise(mean_intensity_0h = mean(Col1_per_cell))
```

### ratio TGF
```{r}
df_ratio_0h <- df_outliersEXnew %>%
  #filter(condition == "TGF") %>%
  left_join(ctrl_df_0h, by = c("plate")) %>%
  mutate(FC_intensity_0h = (Col1_per_cell - mean_intensity_0h) / mean_intensity_0h)%>%
  group_by(time, condition)%>%
  mutate(mean_FC = mean(FC_intensity_0h))

```

#### plot
```{r}
p_test_0h <- df_ratio_0h %>%
 #filter(condition == "TGF") %>%
  ggplot(aes(x = condition, y = FC_intensity_0h)) +
  stat_summary(geom = "crossbar", fun= median, color = "black")+
  ggpubr::stat_compare_means(ref.group = "ctrl", label = "..p.signif..", method = 't.test', hide.ns = T, size = 5, label.y = 7.8, show.legend = F)+# wilcoxon
  ggforce::geom_sina(aes(color = condition, shape = plate),size = 0.8, alpha = 0.6)+
  scale_color_manual(values = c("ctrl" = "#424242", 
                                "TGF" = "#9E9E9E")) +  facet_wrap(~time)+
  facet_wrap(~time)+
  geom_hline(yintercept = 0, color = "gray26", linetype = "longdash") + #C0C0C0 708090 B2BEB5, B0C4DE
  ggtitle("FC = (Col1_per_cell - mean_intensity_0h) / mean_intensity_0h") +
  ylim(-1.5,8.5)+
  theme_bw()
  
p_test_0h

ggplot2::ggsave(filename = paste0(PARAM$folder.figures,
                                  Sys.Date(), "_", "NT22_004_TGFvsctrl0h.png"),
                plot = p_test_0h +
                  theme(strip.text = element_text(size = 10)),#+
                  #theme(axis.text = element_text(size = 11)),
                width = 7, height = 5)
```



